---
title: "Fuel Economoy"
author: "Uzair Ahmad"
date: '2019-12-06'
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{R, load_libraries, message=F, warning=F}
library(sqldf)
library(ggplot2)
```

```{R, echo=FALSE}
vhcls_data = read.csv('https://raw.githubusercontent.com/DrUzair/MLSD/master/Datasets/vehicles.csv')
```

# Asking the right questions: 

## Which manufacturer produces the most fuel efficient fleet of cars?
Considering UCcity be the measure of 'fuel efficiency', 

1. The best UCity vehicle poducer is Hyundai
  - Ioniq Electric- for cosecutive years 2017, 2018, 2019. All three models are EV Front-Wheel Drive vechicles. 

```{r, echo=FALSE}
top_ucity <- head(vhcls_data[order(vhcls_data$UCity, decreasing = TRUE ), c('make', 'model','UCity', 'year')], 10)
print(top_ucity)
```

```{R, echo=FALSE}
year_ucity <- sqldf('select max(Ucity) Best_UCity, make, drive, atvType, fuelType, fuelType1, fuelType2, year from vhcls_data group by year ')
#unique(year_ucity$make)
ggplot(year_ucity, aes(x=year, y=Best_UCity, color=make, shape=drive)) + 
  geom_point(size=2) +
  geom_text(label=substr(as.character(year_ucity[, 'atvType']), 1, 2), size=3, hjust=1, vjust=1)
```

2. Top 10 best UCity vehicles include Tesla, Scion and BMW vehicles.
3. Over the years, the UCity has seen dramatic increase in last decade. All vehicles boasting UCity greater than 100 are EV Rear-wheel Drives except Hyundai's (the winner) Ioniq which is Front-wheel drive.
4. Pre 2000s, Honda vehicles reported best UCity.
5. Interestingly, top 10 best UCity vehicles are Automatic Transmission vehicles.
6. Infact top 217 vehicles, w.r.t UCity, are EV vehicles. Which makes atvType an influential factor to determine UCity.